Fitbit & ET Glasses - Methods

### 2022-09-30 02_paper/02_study/02_fitbit paper/FitBit_Plots_MK.Rmd
upload fitbit data

---
title: "Fitbit & ET Glasses - Methods"
output: html_notebook
---
```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Packages
# install.packages("needs")
# Choose "No"
library(needs)
needs(tidyverse,
      janitor,
      lubridate,
      readxl,
      ggthemes,
      gridExtra,
      imputeTS)
```

@@

Run following Code if and only if new data is added to the initial data.

```{r message=FALSE, warning=FALSE,include=FALSE}
source(file = "data.wrangling.R")
```
# ET Glasses

## Acceleration

The Tobii Glases record acceleration data with a frequency of 100Hz. Therefore it should be possible to calculate the walked distance wearing the glasses using the following formulas:

$$\Delta x_{d,i} = 0.5 (t_{i-1}-t_{i})^2, \text{ with }i=2,...,t_n; d = z,x$$

$$x_d=\sum_{i=1}^n |\Delta x_{d,i}|, \text{ with }i=2,...,t_n; d = z,x$$

$$\sum x=\sqrt{x_z^2+x_x^2}$$

Using these formulas, one brave assumption must be made, that is, for every time period the initial velocity is equal to 0.

The real data gives plausible values.

```{r}
df_meter_real <- read_rds(file = "final_data/meter_real.rds")
df_meter_real
summary(df_meter_real$ges)
```

On the contrary, simulated situations revealed multiple problems.

For the first simulation the participant was standing still without moving its head and with strong movements of the head respectively. A reliable measurement would give the same distance for both situations. 
Afterwards, we simulated two two times 5 meter long walks. One time the participant turned around after 5 meters. During the other walk the participant walked 5 meters and walked back backwards. Again, we expect the glasses to give the same walking distance.
To test real world situations we simulated two random walks in a simulated class room. One walk simulated an extensive walk and the other walk simulated a less intense walk.
Additionally, we tested the values for taking the elevator and walking a staircase.
The simulated situations give the following values.

```{r}
df_meter_sim <- read_rds(file = "final_data/meter_test.rds")

df_meter_sim
```
Plotting the acceleration reveals the problems with the acceleration data.

```{r}
df_meter_sim_time <- read_rds(file = "final_data/meter_test_time.rds")

df_meter_sim_time %>%
  pivot_longer(cols = contains("acc"),
               names_to = "acc") %>%
ggplot(mapping = aes(x = time,
                     y = value
                     )
       ) +
  geom_path() +
  labs(y = "") +
  theme(text = element_text(family = "serif")) +
  facet_grid(rows = vars(name),
             cols = vars(acc) 
             ) +
  theme(strip.text.y = element_text(angle = 0))
```

## Pupil Dialation

The ET Glasses report the pupil dilation at a frequency of 100 Hz.

```{r}
# df_dil <- read_rds(file = "final_data/dilation.rds")
# 
# df_dil %>%
#   pivot_longer(cols = right:left,
#                names_to = "dilation") %>%
# ggplot(mapping = aes(x = time,
#                      y = value
#                      )
#        ) +
#   geom_path() +
#   facet_grid(rows = vars(ID),
#              cols = vars(dilation)
#              )
```

@@

@@

############### FLOW PLOT #############
plot_flow <-
ggplot(data = df_heart_all,
       mapping = aes(x = as.numeric(Time),
                     y = `Heart Rate`,
                     group = ID)
       ) +
  geom_line(mapping = aes(color = status),
            size = 0.3,
            alpha = 1
            ) +
  xlim(0, min(df_heart_min$min)) +
  scale_color_viridis_d(option = "A",
                       end = 0.8)  +
  scale_linetype_manual(values = c(1, 6)) +
  labs(x = "Time (in Seconds)",
       y = "Heart Rate\n(in Beats per Minute)") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.position = "none",
        strip.text = element_text(size = 16)) +
  facet_grid(cols = vars(Type),
             scales = "free_x")

plot_flow_leg <-
ggplot(data = df_heart_all,
       mapping = aes(x = Time,
                     y = `Heart Rate`
                     )
       ) +
  geom_smooth(mapping = aes(color = status),
              method = "loess") +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
 # scale_linetype_manual(values = c(1, 6)) +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.title = element_blank())

@@

@@

plot_loess <-
ggplot(data = filter(df_heart_all),
       mapping = aes(x = as.numeric(Time),
                     y = `Heart Rate`
                     )
       ) +
  stat_smooth(geom = "line",
              mapping = aes(color = status),
              method = "loess",
              se = FALSE) +
  xlim(0,min(df_heart_min$min)) +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  scale_linetype_manual(values = c(1, 6)) +
  labs(x = "Time (in Seconds)",
       y = "Heart Rate\n(in Beats per Minute)") +
  theme_minimal() +
  facet_grid(cols = vars(Type),
             scales = "free_x",
             ) +
  theme(text = element_text(family = "serif"),
        legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

@@

plot_box <- 
ggplot(data = df_heart_aggr,
       mapping = aes(x = Type,
                     y = mean,
                     color = status
                     )
       ) +
  geom_boxplot(color = "darkgrey",
               fill = "grey") +
  geom_line(data = filter(df_heart_aggr,
                          status == "Novice"),
            mapping = aes(x = Type,
                          y = mean,
                          group = as_factor(ID)
                          ),
            linetype = 1,
            alpha = 0.5
            ) +
  geom_line(data = filter(df_heart_aggr,
                          status == "Expert"),
            mapping = aes(x = Type,
                          y = mean,
                          group = as_factor(ID)
                          ),
            linetype = 6,
            alpha = 0.5
            ) +
  geom_point(size = 2,
             alpha = 0.5) +
  labs(x = "",
       y = "Average Heart Rate\n(in Beats per Minute)") +
  scale_color_viridis_d(option = "A",
                       end = 0.8) +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        strip.text = element_blank(),
       legend.position = "none",
        axis.text.x = element_blank()
        ) +
  aes(fill = Type)

lo_mat <- rbind(c(1,1,1,1,1,1,NA),
                c(2,2,2,2,2,2,4),
                c(3,3,3,3,3,3,NA))

plot_heart <- 
grid.arrange(grobs = list(plot_flow, plot_loess, plot_box, legend),
             layout_matrix = lo_mat)

plot_heart
```

>>>>

---
title: "Fitbit & ET Glasses - Methods"
output: html_notebook
---
```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Packages
# install.packages("needs")
# Choose "No"
library(needs)
needs(tidyverse,
      janitor,
      lubridate,
      readxl,
      ggthemes,
      gridExtra,
      imputeTS,
      DescTools,
      cowplot,
      rstatix,
      ggpubr)
```

@@

Run following Code only if new data is added to the initial data.

```{r message=FALSE, warning=FALSE,include=FALSE}
# fitbit data
source(file = "data.wrangling.R")

# demographic data
# prepare data (selected from questionnaire data)
demo.data <- readxl::read_xlsx("./Data/data_empschul_labor_lehrperson.xlsx")

# select relevant columns
demo.data <- demo.data %>% select(LI02_01_1,
                                  LI03_01,
                                  LI04_01,
                                  LI06_05)

# rename columns
demo.data <- rename(demo.data, c("ID" = "LI06_05",
                                  "Gender" = "LI02_01_1",
                                  "Age" = "LI03_01",
                                  "Teaching Experience" = "LI04_01"))

# exclude cases with no fitbit data
demo.data <- demo.data %>% filter(!ID %in% c("126",
                                             "132",
                                             )
                                  )

@@

```{r}
# aggregate data

@@

# Mean, SD and Cohen D for experts & novices
# novices
mean_nov <- 
  df_heart_aggr %>%
  filter(status == "Novice" ) %>% 
  pull(mean) %>% 
  mean() %>% 
  round(., digits = 0)

sd_nov <- 
  df_heart_aggr %>%
  filter(status == "Novice") %>% 
  pull(mean) %>% 
  sd() %>% 
  round(., digits = 0)

# experts
mean_exp <-
  df_heart_aggr %>%
  filter(status == "Expert") %>% 
  pull(mean) %>% 
  mean() %>% 
  round(., digits = 0)

sd_exp <- 
  df_heart_aggr %>%
  filter(status == "Expert") %>% 
  pull(mean) %>% 
  sd() %>% 
  round(., digits = 0)

# effect size for expertise differences
d_heartrate <- CohenD(x = df_heart_aggr$mean[df_heart_aggr$status == "Expert"],
               y = df_heart_aggr$mean[df_heart_aggr$status == "Novice"],
               na.rm = TRUE)

############## FLOW PLOT #############
plot_flow <-
ggplot(data = df_heart_all,
       mapping = aes(x = as.numeric(Time),
                     y = `Heart Rate`,
                     group = ID)
       ) +
  geom_line(mapping = aes(color = status),
            size = 0.3,
            alpha = 1
            ) +
  xlim(0, 600) +
  scale_color_viridis_d(option = "A",
                       end = 0.8)  +
  scale_linetype_manual(values = c(1, 6)) +
  labs(x = "Time (in Seconds)",
       y = "Heart Rate\n(in Beats per Minute)") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.position = "none",
        strip.text = element_text(size = 16)) +
  facet_grid(cols = vars(Type),
             scales = "free_x")

############### FLOW PLOT EXPERT NOVICE #############

plot_flow_leg <-
ggplot(data = df_heart_all,
       mapping = aes(x = Time,
                     y = `Heart Rate`
                     )
       ) +
  xlim(0, 600) +
  geom_smooth(mapping = aes(color = status),
              method = "loess") +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  scale_linetype_manual(values = c(1, 6)) +
  labs(x = "Time (in Seconds)",
       y = "Heart Rate\n(in Beats per Minute)") +
  ggtitle("Experts' and novices' heart rate per phase") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        strip.text = element_text(size = 16),
        plot.title = element_text(size = 20, 
                                  face = "bold")) +
  facet_grid(cols = vars(Type),
             scales = "free_x")

@@

# saving plot 
ggsave(plot = plot_flow_leg,
       filename = "Plots/plot_flow_leg.svg",
       height = 5,
       width = 8,
       units = "in")

@@

plot_loess <-
ggplot(data = filter(df_heart_all),
       mapping = aes(x = as.numeric(Time),
                     y = `Heart Rate`
                     )
       ) +
  stat_smooth(geom = "line",
              mapping = aes(color = status),
              method = "loess",
              se = FALSE) +
  xlim(0,600) +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  scale_linetype_manual(values = c(1, 6)) +
  labs(x = "Time (in Seconds)",
       y = "Heart Rate\n(in Beats per Minute)") +
  theme_minimal() +
  facet_grid(cols = vars(Type),
             scales = "free_x",
             ) +
  theme(text = element_text(family = "serif"),
        legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

@@

##### Box 1 Plot ##### 
plot_box1 <- 
ggplot(data = df_heart_aggr,
       mapping = aes(x = Type,
                     y = mean,
                     color = status
                     )
       ) +
  geom_boxplot(color = "darkgrey",
               fill = "grey") +
  geom_line(data = filter(df_heart_aggr,
                          status == "Novice"),
            mapping = aes(x = Type,
                          y = mean,
                          group = as_factor(ID)
                          ),
            linetype = 1,
            alpha = 0.5
            ) +
  geom_line(data = filter(df_heart_aggr,
                          status == "Expert"),
            mapping = aes(x = Type,
                          y = mean,
                          group = as_factor(ID)
                          ),
            linetype = 6,
            alpha = 0.5
            ) +
  geom_point(size = 2,
             alpha = 0.5) +
  labs(x = "",
       y = "Average Heart Rate\n(in Beats per Minute)") +
  scale_color_viridis_d(option = "A",
                       end = 0.8) +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        strip.text = element_blank(),
        legend.position = "none",
        axis.text.x = element_blank()
        ) +
  aes(fill = Type)

plot_box1

##### Box 2 Plot ##### 
plot_box <-
  ggplot(data = df_heart_aggr,
         mapping = aes(x = Type,
                       y = mean)) + 
  geom_boxplot(mapping = aes(fill = status),
               outlier.shape = NA) +
  geom_point(mapping = aes(shape = status),
             alpha = 0.5,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  labs(x ="",
       y = "Average Heart Rate\n(in Beats per Minute)") + 
  ggtitle("Experts' and novices' heart rate per phase") +
  scale_fill_brewer(palette  = "RdBu") +  
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        strip.text = element_text(size = 16),
        plot.title = element_text(size = 20, 
                                  face = "bold"))

plot_box

# saving plot 
ggsave(plot = plot_box,
       filename = "Plots/plot_box.svg",
       height = 5,
       width = 8,
       units = "in")

##############

# lo_mat <- rbind(c(1,1,1,1,1,1,NA),
#                 c(2,2,2,2,2,2,4),
#                 c(3,3,3,3,3,3,NA))
# 
# plot_heart <- 
# grid.arrange(grobs = list(plot_flow, plot_loess, plot_box1, legend),
#              layout_matrix = lo_mat)
# 
# plot_heart

# Repeated Measures ANOVA

```{r}

# merge demo data and fitbit data
df_anova <- merge(demo.data,
                  df_heart_aggr,
                  by="ID")

# clean variable names
df_anova <- df_anova %>% clean_names()

# check for outliers
outliers <-
  df_anova %>%
  group_by(type) %>%
  identify_outliers(mean)

# check for normality assumption --> can be checked by computing Shapiro-Wilk test for each time point. If the data is normally distributed, the p-value should be greater than 0.05
norm_assump <-
  df_anova %>%
  group_by(type) %>%
  shapiro_test(mean)

ggqqplot(df_anova, "mean", facet.by = "type")

# anova
res.aov <- rstatix::anova_test(data = df_anova,
                               dv = mean,
                               wid = id,
                               within = type)
get_anova_table(res.aov)

# anova expert novice
res.aov2 <- rstatix::anova_test(data = df_anova,
                               dv = mean,
                               between = status,
                               wid = id,
                               within = type)
get_anova_table(res.aov2)

# pairwise comparisons
pwc <- df_anova %>%
  pairwise_t_test(
    mean ~ type, paired = TRUE,
    p.adjust.method = "bonferroni"
    )
pwc

# Visualization: box plots with p-values
pwc <- 
  pwc %>% 
  add_xy_position(x = "type")

# boxplot
bxp_anova <- ggboxplot(data = df_anova, 
                 x = "type", 
                 y = "mean", 
                 add = "point")

bxp_anova + 
  stat_pvalue_manual(pwc) +
  labs(
    subtitle = get_test_label(res.aov, detailed = TRUE),
    caption = get_pwc_label(pwc)
    )

# saving plot 
ggsave(plot = bxp_anova,
       filename = "Plots/bxp_anova.svg",
       height = 5,
       width = 8,
       units = "in")
